
```{r solutions, echo=FALSE}
# This is optional to prevent annoying errors when deploying shiny apps
if(!exists(".flexteach_solutions"))
  .flexteach_solutions = FALSE
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE
  , dev = "svg"
  , dpi = 150
  , fig.width = 5
  , fig.height = 5
  , dev.args = list(pointsize=13)
  , out.width = "40%"
  )
library(broom)
library(gt)
library(dplyr)
library(tidyr)
```

## Exercises

1. Download the data above and import it into SPSS.


```{r}
cn = names(.flexteach_info$data)
```


2. Create a scatterplot.

```{r eval = .flexteach_solutions}
plot(.flexteach_info$data, pch = 19, col = rgb(0,0,0,.3), cex = .8, las = 1)
abline(lm(y ~ x, data = .flexteach_info$data))
```


3. Compute the mean and standard deviation of each of the two variables, ``r cn[1]`` and ``r cn[2]``.

```{r eval = .flexteach_solutions}
.flexteach_info$data %>%
  pivot_longer(c("x","y"), names_to = "variable", values_to = "x") %>%
  group_by(variable) %>%
  summarise(
    N = n(),
    Mean = mean(x),
    SD = sd(x)
    ) -> sum_stats 

sum_stats %>%
  gt(rowname_col = "variable") %>%
  cols_label(SD = "Std. Dev.") %>%
  tab_header(title="Summary statistics") %>%
  fmt_number(c("Mean","SD"), decimals = 3)
```


4. Compute the correlation between the two variables ``r cn[1]`` and ``r cn[2]``.

```{r results = 'asis', eval = .flexteach_solutions}
r = cor(.flexteach_info$data)[2] %>% 
  round(3)
glue::glue("> The correlation is {r}.") %>%
  cat()
```


5. Use the summary statistics and correlation you computed above to compute the slope of the regression line.

```{r eval = .flexteach_solutions, child='children/slope.Rmd'}
```


6. Use the summary statistics, correlation, and slope you computed above to compute the intercept of the regression line. Use ``r cn[1]`` as the independent variable and ``r cn[2]`` as the dependent variable.

```{r eval = .flexteach_solutions, child='children/intercept.Rmd'}
```



7. Perform a linear regression analysis in SPSS.

```{r eval = .flexteach_solutions}

lm_obj = lm(y ~ x, data = .flexteach_info$data)
r2 = summary(lm_obj)$r.square
ar2 = summary(lm_obj)$adj.r.square

s = summary(lm_obj)$sigma
preds = glue::glue("Predictors: {paste( names(coef(lm_obj)), collapse=', ')}")

data.frame(
  Model = "1", 
  R = sqrt(r2),
  r2 = r2,
  ar2 = ar2,
  s = s
) %>%
  gt() %>%
  cols_label(
    r2 = "R Square",
    ar2 = "Adjusted R Square",
    s = "Std. Error of the Estimate"
  ) %>%
  tab_footnote(preds, cells_body(columns = "R")) %>%
  fmt_number(c("R","r2","ar2","s"), decimals = 3) %>%
  tab_header("Model summary")

```

<br/>
<br/>


```{r eval = .flexteach_solutions}

std_beta = c(NA, QuantPsyc::lm.beta(lm_obj))

lm_obj %>%
  tidy() %>%
  mutate(std_beta = std_beta) %>%
  relocate(std_beta, .before = "statistic") %>%
  gt() %>%
  cols_label(
    term="",
    estimate="B",
    std_beta = "Beta",
    std.error="Std. Error",
    statistic="t",
    p.value="Sig."
  ) %>%
  fmt_missing(everything(),missing_text = "") %>%
  fmt_number(
    c("std.error","std_beta","statistic","p.value"),
    decimals=3
  ) %>%
  text_transform(
    locations = cells_body(columns = "estimate"),
    fn = function(x){
      x = as.numeric(x)
      z = ifelse(abs(x)>=.0005,
                 round(x, 3),
                 toupper(format(x,digits=4,scientific=TRUE))
      )
      as.character(z)
    }) %>%
  tab_header("Coefficients") %>%
  tab_spanner("Unstandardized", c("estimate","std.error")) %>%
  tab_spanner("Standardized", "std_beta")

```

```{r results = 'asis', eval = .flexteach_solutions}
cat(
"
> Note that the values given by SPSS may be slightly different than the ones you calculated due to the rounding in the calculation. When you calculate it, it is best not to round until the end (unlike what is shown in questions 5 and 6, where values were rounded after each step for neatness of the display).
")
```

