---
title: "Summary statistics"
author: "Richard D. Morey"
date: "22 September 2016"
output:
  html_document:
      dev: svg
  html_fragment:
      dev: svg
---

```{r include=FALSE}
# This chunk allows standalone compilation
if(!exists("assignment")){
  seed = 124
  secret = "dgnvo"
  source("assignment.R")
  source("../../settings.R")
  solutions = FALSE
}

```

`r assignment`
-----

```{r include=FALSE}
knitr::opts_chunk$set(fig.width=5,fig.height=5)

require(xtable)
x = getData(seed, secret, assignment)$x
N = length(x)

```


Suppose you have a sample of $N=`r N`$ random values, and you'd like to compute the summary statistics for the sample.

```{r echo=FALSE, results = 'asis'}
    res = matrix(x,nrow=1)
    colnames(res) = paste("$x_{",1:N,"}$", sep="")
    knitr::kable(res, digits = 4, format = "html", table.attr = "class=\"anovatab\"")
```

First, we compute the mean.

## Compute the sample mean

There are three steps to computing a sample mean: Sum all the values, divide the sum by the total number of elements, then interpret the result.

### 1. Compute the sum

First, add up all the values.

```{r echo=FALSE,eval=solutions,child="child_sum.Rmd"}
```

### 2. Divide by the number of elements

Now take the sum you computed in the last step, and divide it by $N$, the total number of elements.

```{r echo=FALSE,eval=solutions,child="child_divideN.Rmd"}
```

### 3. Interpret the result

The mean is has several interpretations. It is:

* Informally, a value in the "middle" of the data values.
* The value for which the sum of the deviations from that value is exactly 0.
* The value that minimizes the squared deviations from it.

```{r echo=FALSE,eval=solutions,child="child_plotMean.Rmd"}
```


