---
title: "Computing residuals in SPSS"
author: "Richard D. Morey"
date: "22 September 2016"
output:
  html_document:
      dev: svg
  html_fragment:
      dev: svg
---

```{r include=FALSE}
# This chunk allows standalone compilation
if(!exists("assignment")){
  seed = 124
  secret = "dgnvo"
  source("assignment.R")
  source("../../settings.R")
  solutions = FALSE
}

```

`r assignment`
-----


```{r include=FALSE}
knitr::opts_chunk$set(fig.width=5,fig.height=5)

require(stargazer)
require(broom)
all.data = getData(seed, secret, assignment)
myData = all.data[['data']]
infoData = all.data[['info']]
```


The (fictional) data set for this assignment has two columns, `x` and `y`. 

## Part 1: Linear regression

Load the data set into SPSS. Create a scatterplot with `x` and `y` as the independent and dependent variables respectively. Describe the pattern you see and choose one of the following measures of the strength of the correlation:

* Pearson's $r$
* Kendall's $\tau$
* None (correlation is inappropriate)

*Regardless of whether linear correlation is appropriate*, use SPSS to add the least squares linear regression line.

### Step by step in SPSS


```{r echo=FALSE,fig.cap="Scatterplot",include=solutions}

par(las=1,mar=c(3.5,3.5,4,.5),mgp=c(2.5,1,0))

x = myData$x
y = myData$y

res = lm(y~x)

plot(x, y, ylab="y",xlab="x",pch=1, main="Scatterplot")
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = 
rgb(0,0,0,.05))
points(x,y)
abline(res)
```

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution\n",
    infoData$text)
```


## Part 2: residuals

```{r echo=FALSE,fig.cap="Residuals",include=solutions}

par(las=1,mar=c(3.5,3.5,4,.5),mgp=c(2.5,1,0))

plot(x, res$residuals, ylab="Residuals",xlab="x",pch=1, main="Residuals")
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = 
rgb(0,0,0,.05))
points(x,res$residuals)
abline(h=0,lty=2)
lines(lowess(res$residuals~x), col="red", lty=2)
```

### Step by step in SPSS


```{r include=FALSE, eval=FALSE}
# This is a test of the plotly engine, which doesn't seem to work
# right now due to this being in an html_fragment
library(plotly)
plot_ly(iris, x = ~Sepal.Length, y = ~Petal.Length)
```


