---
title: "Regression assignment"
author: "Richard D. Morey"
date: "8 November 2016"
output:
  html_document:
      dev: svg
  html_fragment:
      dev: svg
---

```{r include=FALSE}
# This chunk allows standalone compilation
if(!exists("assignment")){
  seed = 124
  secret = "dgnvo"
  source("assignment.R")
  source("../../settings.R")
  solutions = FALSE
}

```

`r assignment`
-----

## Instructions

This is your second assignment for PS2024.  Please type your answers and/or copy/paste the appropriate output from SPSS into a single document so that your assignment can be submitted electronically.

To do this assignment you can use the Field textbook (3rd Edition, chapter 12 or 4th Edition, chapter 13), the first part of Chapter 13 from Howell, other books or the class notes.  


Consult the assignment sheet for the first assignment to find out how to use this website. Please download the data using either the <a href="https://www.mozilla.org/en-US/firefox/new/" target="_blank">Mozilla Firefox</a> or the <a href="https://www.google.com/chrome/" target="_blank">Google Chrome</a> web browser. 


```{r include=FALSE}
knitr::opts_chunk$set(fig.width=5,fig.height=5)

require(stargazer)
require(broom)
all.data = getData(seed, secret, assignment)
myData = all.data[['data']]
infoData = all.data[['info']]

```


The (fictional) data set for this assignment has ten columns. The names, along with their descriptions, appear in the table below.

Variable name | Description
--------------|----------------
`ID` | Person identifier 
`stress.before.test1` | Self-rated stress before a particular vocabulary test 
`score.test1` |  Score on a particular vocabulary test
`IQ` | Intelligence quotient
`cognitive.task2` | Accuracy on a particular cognitive task
`practice.task2` | Amount of practice on cognitive task in days
`response.time.task2` | Average response time on cognitive task
`college.math` | Score on a standardized college entrance math exam
`score.math.course1` | Mark in first high school math course      
`score.math.course2` | Mark in second high school math course        

## Question 1

Load the data set into SPSS. For each of the pairs of variables below, create a scatterplot, choosing the correct dependent and independent variables for the axes. Describe the pattern you see in each of the scatterplots, and for each one, choose one of the following measures of the strength of the correlation:

* Pearson's $r$
* Kendall's $\tau$
* None (correlation is inappropriate)

### Relationship A (1 point)

Suppose we are interested in predicting a student's score on their vocabulary test from the stress they experienced before that test. Describe the relationship in terms of form, strength, and direction (if applicable). Calculate the appropriate correlation coefficient (if any).


```{r echo=FALSE,fig.cap="Relationship A",include=solutions}

par(las=1,mar=c(3.5,3.5,4,.5),mgp=c(2.5,1,0))

x = myData$stress.before.test1
y = myData$score.test1

plot(x, y, ylab="score.test1",xlab="stress.before.test1",pch=1, main="Relationship A")
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = 
rgb(0,0,0,.05))
points(x,y)
```

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution, Relationship A\n",
    infoData$texts[1])
```

### Relationship B (1 point)

Suppose we are interested in predicting a student's score on the cognitive task from their IQ. Describe the relationship in terms of form, strength, and direction (if applicable). Calculate the appropriate correlation coefficient (if any).

```{r echo=FALSE,fig.cap="Relationship B",include=solutions}

par(las=1,mar=c(3.5,3.5,4,.5),mgp=c(2.5,1,0))

x = myData$IQ
y = myData$cognitive.task2

plot(x, y, ylab="cognitive.task2",xlab="IQ",pch=1, main="Relationship B")
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = 
rgb(0,0,0,.05))
points(x,y)
```

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution, Relationship B\n",
    infoData$texts[2])
```

### Relationship C (1 point)

Suppose we are interested in predicting a student's response time on the cognitive task from the amount of time they practiced on the cognitive task. Describe the relationship in terms of form, strength, and direction (if applicable). Calculate the appropriate correlation coefficient (if any).

```{r echo=FALSE,fig.cap="Relationship C",include=solutions}

par(las=1,mar=c(3.5,3.5,4,.5),mgp=c(2.5,1,0))

x = myData$practice.task2
y = myData$response.time.task2

plot(x, y, ylab="response.time.task2",xlab="practice.task2",pch=1, main="Relationship C")
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = 
rgb(0,0,0,.05))
points(x,y)
```

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution, Relationship C\n",
    infoData$texts[3])
```


## Question 2 

Perform a linear regression analysis, regressing the college entrance exam math score onto the second high school math course score (that is, the `score.math.course2` will be the independent variable).

### a) (1 point)

Create a scatterplot. 

```{r echo=FALSE,results='asis',include=solutions}
cat("#### Solution\n\n")
```

```{r echo=FALSE,include=solutions}
par(las=1,mar=c(3.5,3.5,4,.5),mgp=c(2.5,1,0))

x = myData$score.math.course2
y = myData$college.math

plot(x, y, ylab="score.test1",xlab="score.math.course2",pch=1, main="Entrance score from second course mark")
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = 
rgb(0,0,0,.05))
points(x,y)
#abline(lm(y~x))

```

### b) (2 points)

How well does the `score.math.course2` predict scores on the college entrance exam? Report a Pearson $r$ or Kendall's $\tau$ correlation, as appropriate, as well as a $p$ value on the slope of the relationship.

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution
The typical correlation coefficient one would use here is Pearson's $r$, which is $r=",round(cor(x,y),3),"$. Kendall's $\\tau$ is also possible; $\\tau= ",round(cor(x,y,method="kendall"),3),"$.",
sep="")
```

```{r echo=FALSE,results="asis",include=solutions}
    res = lm(college.math ~ score.math.course2, data=myData)
    knitr::kable(tidy(res),digits = 4, format = "html", table.attr = "class=\"anovatab\"")
```

## Question 3 

Suppose we which to try to predict peoples' college entrance scores more accurately, and we add `score.math.course1` as a second predictor (independent variable). 

### a) (2 points)

First perform the linear regression analysis with only `score.math.course1` as the independent variable, then perform the linear regression analysis with *both* high school math courses as predictors.

```{r echo=FALSE, results='asis', echo=FALSE,include=solutions}
res = lm(college.math ~ score.math.course1, dat=myData)

cat("#### Solution\n\n")

    knitr::kable(tidy(res),digits = 4, format = "html", table.attr = "class=\"anovatab\"")
```

```{r echo=FALSE}
res = lm(college.math ~ score.math.course1 + score.math.course2, dat=myData)
```

Report $p$ values for both slopes. 

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution\n\n")

    knitr::kable(tidy(res),digits = 4, format = "html", table.attr = "class=\"anovatab\"")
```

### b) (1 point)

Report $R^2$ and adjusted $R^2$. 

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution\n")

stargazer::stargazer(res, type = 'html')


```

### c) (1 point)

Explain the behaviour of the $p$ values in the linear regression analysis above.

```{r echo=FALSE, results='asis',include=solutions}

cat("#### Solution\n\n Write several sentences. Your solution will be evaluated by the marker. ")

```

