---
title: "Regression assignment"
author: "Richard D. Morey"
date: "13 November 2017"
output:
  html_document:
      dev: svg
  html_fragment:
      dev: svg
---

```{r include=FALSE}
# This chunk allows standalone compilation
if(!exists("assignment")){
  seed = 124
  secret = "dgnvo"
  source("assignment.R")
  source("../../settings.R")
  solutions = FALSE
}

getData2 <- function(seed, secret, extra = ""){
  
  set.seed.alpha(paste0(seed, secret, extra))

  n = 4
  y = 1:n
  x = sample(y)
  
  data.frame(x = jitter(x), y = jitter(y))  
  
}

```

`r assignment`
-----

## Instructions

This is your second assignment for PS2024.  Please type your answers and/or copy/paste the appropriate output from SPSS into a single document so that your assignment can be submitted electronically.

To do this assignment you can use the Field textbook (3rd Edition, chapter 12 or 4th Edition, chapter 13), the first part of Chapter 13 from Howell, other books or the class notes.  

Consult the assignment sheet for the first assignment to find out how to use this website. 

```{r include=FALSE}
knitr::opts_chunk$set(fig.width=5,fig.height=5)

require(stargazer)
require(broom)
all.data = getData(seed, secret, assignment)
myData = all.data[['data']]
infoData = all.data[['info']]

```


The (fictional) data set for this assignment has ten columns. The names, along with their descriptions, appear in the table below.

Variable name | Description
--------------|----------------
`ID` | Person identifier 
`stress.before.test1` | Self-rated stress before a particular vocabulary test 
`score.test1` |  Score on a particular vocabulary test
`IQ` | Intelligence quotient
`cognitive.task2` | Accuracy on a particular cognitive task
`practice.task2` | Amount of practice on cognitive task in days
`response.time.task2` | Average response time on cognitive task
`college.math` | Score on a standardized college entrance math exam
`score.math.course1` | Mark in first high school math course      
`score.math.course2` | Mark in second high school math course        

## Question 1

Load the data set into SPSS. For each of the pairs of variables below, create a scatterplot, choosing the correct dependent and independent variables for the axes. Describe the pattern you see in each of the scatterplots, and for each one, choose one of the following measures of the strength of the correlation:

* Pearson's $r$
* Kendall's $\tau$
* None (correlation is inappropriate)

### Relationship A (1 point)

Suppose we are interested in predicting a student's score on their vocabulary test from the stress they experienced before that test. Describe the relationship in terms of form, strength, and direction (if applicable). Calculate the appropriate correlation coefficient (if any).


```{r echo=FALSE,fig.cap="Relationship A",include=solutions}

par(las=1,mar=c(3.5,3.5,4,.5),mgp=c(2.5,1,0))

x = myData$stress.before.test1
y = myData$score.test1

plot(x, y, ylab="score.test1",xlab="stress.before.test1",pch=1, main="Relationship A")
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = 
rgb(0,0,0,.05))
points(x,y)
```

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution, Relationship A\n",
    infoData$texts[1])
```

### Relationship B (1 point)

Suppose we are interested in predicting a student's score on the cognitive task from their IQ. Describe the relationship in terms of form, strength, and direction (if applicable). Calculate the appropriate correlation coefficient (if any).

```{r echo=FALSE,fig.cap="Relationship B",include=solutions}

par(las=1,mar=c(3.5,3.5,4,.5),mgp=c(2.5,1,0))

x = myData$IQ
y = myData$cognitive.task2

plot(x, y, ylab="cognitive.task2",xlab="IQ",pch=1, main="Relationship B")
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = 
rgb(0,0,0,.05))
points(x,y)
```

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution, Relationship B\n",
    infoData$texts[2])
```

### Relationship C (1 point)

Suppose we are interested in predicting a student's response time on the cognitive task from the amount of time they practiced on the cognitive task. Describe the relationship in terms of form, strength, and direction (if applicable). Calculate the appropriate correlation coefficient (if any).

```{r echo=FALSE,fig.cap="Relationship C",include=solutions}

par(las=1,mar=c(3.5,3.5,4,.5),mgp=c(2.5,1,0))

x = myData$practice.task2
y = myData$response.time.task2

plot(x, y, ylab="response.time.task2",xlab="practice.task2",pch=1, main="Relationship C")
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = 
rgb(0,0,0,.05))
points(x,y)
```

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution, Relationship C\n",
    infoData$texts[3])
```


```{r child='oldQ2-3.Rmd', eval=FALSE}
```

## Question 2

**Make sure you've put in your correct student ID in the ID box for the marked version of the assignment! The scatterplot below depends on your ID number.** 

Consider the scatterplot below.

```{r echo=FALSE,results='asis'}
tab = getData2(seed, secret,"")
plot(tab$x,tab$y,xlab = "X", ylab = "Y", pch = 19, col = "blue", las = 1)
```

### a) (1 point)

Count the number of total pairs of points, the number of pairs with a showing a positive (consistent), and the number of pairs showinf a negative (inconsistent) relationship.

```{r echo=FALSE, child='kendall1.Rmd',eval=solutions}
```


### b) (2 points)

Use the numbers computed in part (a) to calculate the proportion of positive (consistent) pairs, and then compute Kendall's $\tau$.

```{r echo=FALSE, child='kendall2.Rmd',eval=solutions}
```

## Question 3

Use what you know about how Kendall's $\tau$ and Pearson's $r$ are computed and interpreted to answer the following. It might help to consider the scatterplot from Question 2, and how $\tau$ and $r$ would change if you moved various points in the scatterplot.

### a) (2 points)

Explain, in your own words, how Pearson's $r$ can indicate a *strong* relationship while Kendall's $\tau$ indicates a *weak* relationship.

### b) (2 points)

Explain, in your own words, how Pearson's $r$ can indicate a *weak* relationship while Kendall's $\tau$ indicates a *strong* relationship.




