## Question 2 

Perform a linear regression analysis, regressing the college entrance exam math score onto the second high school math course score (that is, the `score.math.course2` will be the independent variable).

### a) (1 point)

Create a scatterplot. 

```{r echo=FALSE,results='asis',include=solutions}
cat("#### Solution\n\n")
```

```{r echo=FALSE,include=solutions}
par(las=1,mar=c(3.5,3.5,4,.5),mgp=c(2.5,1,0))

x = myData$score.math.course2
y = myData$college.math

plot(x, y, ylab="score.test1",xlab="score.math.course2",pch=1, main="Entrance score from second course mark")
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = 
rgb(0,0,0,.05))
points(x,y)
#abline(lm(y~x))

```

### b) (2 points)

How well does the `score.math.course2` predict scores on the college entrance exam? Report a Pearson $r$ or Kendall's $\tau$ correlation, as appropriate, as well as a $p$ value on the slope of the relationship.

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution
The typical correlation coefficient one would use here is Pearson's $r$, which is $r=",round(cor(x,y),3),"$. Kendall's $\\tau$ is also possible; $\\tau= ",round(cor(x,y,method="kendall"),3),"$.",
sep="")
```

```{r echo=FALSE,results="asis",include=solutions}
    res = lm(college.math ~ score.math.course2, data=myData)
    knitr::kable(tidy(res),digits = 4, format = "html", table.attr = "class=\"anovatab\"")
```

## Question 3 

Suppose we which to try to predict peoples' college entrance scores more accurately, and we add `score.math.course1` as a second predictor (independent variable). 

### a) (2 points)

First perform the linear regression analysis with only `score.math.course1` as the independent variable, then perform the linear regression analysis with *both* high school math courses as predictors.

```{r echo=FALSE, results='asis', echo=FALSE,include=solutions}
res = lm(college.math ~ score.math.course1, dat=myData)

cat("#### Solution\n\n")

    knitr::kable(tidy(res),digits = 4, format = "html", table.attr = "class=\"anovatab\"")
```

```{r echo=FALSE}
res = lm(college.math ~ score.math.course1 + score.math.course2, dat=myData)
```

Report $p$ values for both slopes. 

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution\n\n")

    knitr::kable(tidy(res),digits = 4, format = "html", table.attr = "class=\"anovatab\"")
```

### b) (1 point)

Report $R^2$ and adjusted $R^2$. 

```{r results='asis',echo=FALSE,include=solutions}
cat("#### Solution\n")

stargazer::stargazer(res, type = 'html')


```

### c) (1 point)

Explain the behaviour of the $p$ values in the linear regression analysis above.

```{r echo=FALSE, results='asis',include=solutions}

cat("#### Solution\n\n Write several sentences. Your solution will be evaluated by the marker. ")

```

